Parsing Tweets into Universal Dependencies

### 2018-03-04 paper.tex
copyedits & comments

\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
%\newcommand{\yjcomment}[1]{}
%\newcommand{\nascomment}[1]{}
%\newcommand{\yicomment}[1]{}

@@

Universal Dependencies \citep[UD]{NIVRE16.348} were introduced to enable
consistent annotation across different languages. To allow such
consistency, UD was designed to be adaptable to different genres \cite{wang-EtAl:2017:Long6} 
and languages \cite{guo-EtAl:2015:ACL-IJCNLP2,TACL892}. We propose that analyzing
the syntax of tweets can benefit from such adaptability. In this paper,
we introduce a new tweet treebank of 55,607 tokens that follows the UD
guidelines, but also contends with social media-specific challenges that were not
covered by UD guidelines. Our annotation includes 
tokenization, part-of-speech (POS) tags, and (labeled) Universal Dependencies.
We characterize the disagreements among our annotators and find that
consistent annotation is still challenging to deliver even with
the extended guidelines.

@@

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrr}
		\hline
		& train & dev & test \\
		\hline
		tweets 		&	1,639 		&	709		&	1,201\\
		& +1,000 & +709 & +1,000\\
		tokens 		&	24,753		&	11,742	&	19,112\\
		\hline
	\end{tabular}
\caption{Statistics of {\sc Tweebank v2}.
The second row shows the number of new tweets compared to
{\sc Tweebank v1}.\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.
Newly created annotations are splitted into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

We report the inter-annotator agreement between the annotators
in the second stage. There is very little disagreement on the
tokenization annotation. % \nascomment{still better to quantify}.
The agreement on POS is 96.6\%, the unlabeled
dependency agreement is 88.8\% and the labeled dependency agreement
is 84.3\%. Further analysis shows the major disagreements on POS
involve entity names (30.6\%) %\yicomment{noun phrases or proper noun
                              %phrases, do we need example
                              %here?}\yjcomment{i wasn't counting
                              %PROPN and NOUN confusion, but any type
                              %of confusion related with PROPN}
                              %\nascomment{probably don't need examples}
and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' 
%\yicomment{need to relate to previous error analysis, entity names
%maybe?}
can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. 
%\yicomment{need to say the reason why dependency inter-annotator
%agreement is low? 1. POS disagreement lead to big difference in
%understanding the semantics. 2. Some special constructions like
%ellipsis lead to different view of tweets. Fig3 is an example of 2.}
%\nascomment{I don't think we have room to go deeper}
An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of an empty auxiliary verb, or a clause-modified
noun, either annotation is plausible.

@@

\section{Pipeline}
\label{sec:parsing}

@@

\subsection{Tokenizer} \label{sec:tok}
Tokenization, as the initial step of many NLP tasks, is non-trivial for
informal tweets, which include hashtags, at-mentions, and emoticons
\cite{ICWSM101540}.  Context is often required for tokenization
decisions; for example,
the asterisk in {\it 4*3} is a separate token signifying
multiplication, but for the asterisk in {\it sh*t}
works as a mask to evoke censorship and should not be segmented.

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth,trim={0 0 11cm 9cm},clip]{graphics/bilstm_tokenizer}
	\caption{The bi-LSTM tokenizer that segments `{\it its gonna be}' into `{\it it s gon na be}'.}\label{fig:tok-model}
\end{figure}
We introduce a new
character-level bidirectional LSTM (bi-LSTM) sequence-labeling model
\cite{DBLP:journals/corr/HuangXY15,ma-hovy:2016:P16-1}
for tokenization.
Our model takes the raw sentence and tags each character in this 
sentence as whether it is the beginning of a word (1 as the beginning and 0 otherwise).
Figure \ref{fig:tok-model} shows our tokenization model.
Space is treated as an input but deterministically  assigned a special tag \$.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UD pipe v1.2 & 94.7 \\
		\hdashline
		UD pipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser trained 
on the combination of UD\_English and English dialect (Singlish) treebank
outperform the one only on the dialect treebank.  Preliminary results
showed the same in our case, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set.% \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} systems are rule-based systems and were not adapted
to the UD tokenization scheme.
The UD pipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with the UD pipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 against 20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & Precision \\
		\hline
%		Stanford Postagger & 65.6 \\
		Stanford CoreNLP & 90.4 \\
%		UD pipe v1.2 & 78.7 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (greedy) & 94.2 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (CRF) & 95.1 \\
		\hdashline
%		UD pipe v1.2 & 83.8 \\
		\citealp{ma-hovy:2016:P16-1} & 91.8 \\
%		our word bi-LSTM & 89.2 \\
%		our character + word bi-LSTMs & 91.5 \\
		\hline
	\end{tabular}
	\caption{POS tagger comparison on gold-standard tokens in the
          \textsc{Tweebank v2} test set. \label{tbl:pos-result}}
\end{table}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a pipeline of the state-of-the-art systems (labeled  ``SOTA''):
Stanford CoreNLP tokenizer\footnote{We choose Stanford CoreNLP tokenizer in the spirit of comparing rule-based and statistical model-based method.}, %\yicomment{why not compare with UD pipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{dozat-qi-manning:2017:K17-3}'s parser.
Our system differs from the SOTA pipeline in the tokenization and parser components.
From Table \ref{tbl:pipline}, our pipeline outperforms the SOTA when
evaluated in pipeline manner.
The results also emphasize the
importance of tokenization:  without gold tokenization
%\nascomment{changed ``segmentation'' to ``tokenization'' here, twice},
UD parsing
performance drops by more than four points.

>>>>

\newcommand{\yjcomment}[1]{\textcolor{orange}{[$_\mathrm{L}^\mathrm{Y}$#1]}}
\newcommand{\nascomment}[1]{\textcolor{blue}{[#1 ---\textsc{nas}]}}
\newcommand{\yicomment}[1]{\textcolor{gray}{[#1 ---\textsc{yi}]}}
\newcommand{\nss}[1]{\textcolor{magenta}{[$_\mathrm{S}^\mathrm{NS}$#1]}}
%\newcommand{\yjcomment}[1]{}
%\newcommand{\nascomment}[1]{}
%\newcommand{\yicomment}[1]{}

@@

Universal Dependencies \citep[UD]{NIVRE16.348} were introduced to enable
consistent annotation across different languages. To allow such
consistency, UD was designed to be adaptable to different genres \cite{wang-EtAl:2017:Long6} 
and languages \cite{guo-EtAl:2015:ACL-IJCNLP2,TACL892}. We propose that analyzing
the syntax of tweets can benefit from such adaptability. In this paper,
we introduce a new English tweet treebank of 55,607 tokens that follows the UD
guidelines, but also contends with social media-specific challenges that were not
covered by UD guidelines.\footnote{We developed our treebank independently of a similar effort for Italian tweets \cite{sanguinetti-17}. 
See \nss{section ref} for a comparison.} Our annotation includes 
tokenization, part-of-speech (POS) tags, and (labeled) Universal Dependencies.
We characterize the disagreements among our annotators and find that
consistent annotation is still challenging to deliver even with
the extended guidelines.

@@

\paragraph{Comparison to PoSTWITA-UD.} \nss{discuss similarities and differences of conventions to \cite{sanguinetti-17}. 
We expect that the official UD guidelines will eventually adopt standards for these issues so the treebanks can be harmonized.}

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrr}
		\hline
		& train & dev & test \\
		\hline
		tweets 		&	1,639 		&	709		&	1,201\\
		& +1,000 & +709 & +1,000\\
		tokens 		&	24,753		&	11,742	&	19,112\\
		\hline
	\end{tabular}
\caption{Statistics of {\sc Tweebank v2}.
The second row shows the number of new tweets compared to
{\sc Tweebank v1}.\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.\nss{What tools did annotators use?}
Newly created annotations are split into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

We report the inter-annotator agreement between the annotators
in the second stage. There is very little disagreement on the
tokenization annotation. % \nascomment{still better to quantify}.
The agreement on POS is 96.6\%, the unlabeled
dependency agreement is 88.8\% and the labeled dependency agreement
is 84.3\%. Further analysis shows the major disagreements on POS
involve entity names (30.6\%) %\yicomment{noun phrases or proper noun
                              %phrases, do we need example
                              %here?}\yjcomment{i wasn't counting
                              %PROPN and NOUN confusion, but any type
                              %of confusion related with PROPN}
                              %\nascomment{probably don't need examples}
and topical hashtags (18.1\%).
Taking the example in Figure \ref{fig:non-syn-toks}, ``Fix you'' 
%\yicomment{need to relate to previous error analysis, entity names
%maybe?}
can
be understood as a verbal phrase but also as the name of the Coldplay's
single and tagged as proper noun. 
%\yicomment{need to say the reason why dependency inter-annotator
%agreement is low? 1. POS disagreement lead to big difference in
%understanding the semantics. 2. Some special constructions like
%ellipsis lead to different view of tweets. Fig3 is an example of 2.}
%\nascomment{I don't think we have room to go deeper}
An example of a disagreement on
dependencies is shown in Figure \ref{fig:disagree}.  Depending on
whether this is an example of a zero copula construction, or a clause-modified
noun, either annotation is plausible.

@@

\section{Parsing Pipeline}
\label{sec:parsing}

@@

\subsection{Tokenizer} \label{sec:tok}
Tokenization, as the initial step of many NLP tasks, is non-trivial for
informal tweets, which include hashtags, at-mentions, and emoticons
\cite{ICWSM101540}.  Context is often required for tokenization
decisions; for example,
the asterisk in {\it 4*3} is a separate token signifying
multiplication, but the asterisk in {\it sh*t}
works as a mask to evoke censorship and should not be segmented.

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth,trim={0 0 11cm 9cm},clip]{graphics/bilstm_tokenizer}
	\caption{The bi-LSTM tokenizer that segments `{\it its gonna be}' into `{\it it s gon na be}'.}\label{fig:tok-model}
\end{figure}
We introduce a new
character-level bidirectional LSTM (bi-LSTM) sequence-labeling model
\cite{DBLP:journals/corr/HuangXY15,ma-hovy:2016:P16-1}
for tokenization.
Our model takes the raw sentence and tags each character in this 
sentence as whether it is the beginning of a word (1 as the beginning and 0 otherwise).
Figure \ref{fig:tok-model} shows the architecture of our tokenization model.
Space is treated as an input but deterministically  assigned a special tag \$.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?} trained 
on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
outperforms the one trained only on the dialect treebank.  Preliminary results
showed the same in our case, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 vs.~20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.\nss{this paragraph needs copyediting}

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & Precision \\
		\hline
%		Stanford Postagger & 65.6 \\
		Stanford CoreNLP & 90.4 \\
%		UDPipe v1.2 & 78.7 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (greedy) & 94.2 \\
		\citealp{owoputi-EtAl:2013:NAACL-HLT} (CRF) & 95.1 \\
		\hdashline
%		UDPipe v1.2 & 83.8 \\
		\citealp{ma-hovy:2016:P16-1} & 91.8 \\
%		our word bi-LSTM & 89.2 \\
%		our character + word bi-LSTMs & 91.5 \\
		\hline
	\end{tabular}
	\caption{POS tagger comparison on gold-standard tokens in the
          \textsc{Tweebank v2} test set. \label{tbl:pos-result}}
\end{table}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a pipeline of the state-of-the-art systems (labeled  ``SOTA''):
Stanford CoreNLP tokenizer\footnote{We choose Stanford CoreNLP tokenizer in the spirit of comparing rule-based and statistical model-based method.}, %\yicomment{why not compare with UDPipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{dozat-qi-manning:2017:K17-3}'s parser.
Our system differs from the SOTA pipeline in the tokenization and parser components.
From Table \ref{tbl:pipline}, our pipeline outperforms the SOTA when
evaluated in pipeline manner.
The results also emphasize the
importance of tokenization:  without gold tokenization
%\nascomment{changed ``segmentation'' to ``tokenization'' here, twice},
UD parsing
performance drops by more than four points.
